External Patch Group Prior Guided Internal Orthogonal Dictionary Learning for Real Image Denoising

### 2016-10-22 CVPR2017_Review_Unified.tex
add

Though each PG has been projected into its most suitable subspace, the pre-learned subspace is still too general to represent the noisy PG extracted from the real noisy image. That is, the noisy PGs projected into one cluster can still constisted a subspace which is of lower dimensions than the subspace pre-learned from the external PGs. This can be demonstrated by compare the distribution of external PGs and internal PGs in the same clusters. We randomly select one cluster, and collect the celan PGs extracted from external dataset (Kodak 24 images) and the niosy PGs from the testing image. Since the original PGs are of $3p^{2}$ dimensions, we apply PCA to project the PGs into 2 dimensions for better visualization. The results is shown in Figure \ref{fig}, from which we can see clearly that the projected PGs are mainly in a smaller region of the external PGs, which proves that the internal PGs are only consisted a subspace in a lower dimension than the PGs collected from external subspace. To better and adaptively charactering the internal PGs from the testing image, we need learn a more specific dictionary for noisy PGs assigned into each cluster. For notation simplicity, we ignore the index of subspace $k$. The internal PGs $\mathbf{Y}$ form a subspace which can be obtained by singular value decomposition (SVD),
\begin{equation}\label{equ4}
\begin{split}
&\min_{\mathbf{D}\in \mathbb{R}^{3p^2\times r},\textbf{A}\in\mathbb{R}^{3p^2\times MN}}
\|\mathbf{Y}-[\mathbf{D}_{e};\mathbf{D}_{i}]\mathbf{A}\|_{F}^{2}+\lambda\|\mathbf{A}\|_{1}
\\
&
\text{s.t.}
\quad
\mathbf{D}_{i}^{T}\mathbf{D}_{i} = \mathbf{I}_{r},\ \mathbf{D}_{e}^{T}\mathbf{D}_{i} = \mathbf{0},
\end{split}
\end{equation}
The singular vectors capture the statistical structures of NSS variations in natural images, while the singular values in $\mathbf{S}$ represent the significance of these singular vectors. Fig.\ 4 shows the singular vectors for one Gaussian component.

@@

\textbf{Updating Weighted Sparse Coding}: given the initialization orthogonal dictioanry $\textbf{D}_{i}^{(t)}$, the sparce coefficients 
$\textbf{A}^{(t)}$ are obtained via solving
\begin{equation}
\textbf{A}^{(t)}:=\argmin_{\textbf{A}\in\mathbb{R}^{3p^2\times MN}}
\|\mathbf{Y}-[\mathbf{D}_{e};\mathbf{D}_{i}^{(t)}]\mathbf{A}\|_{F}^{2}+\lambda\|\mathbf{A}\|_{1}
\end{equation}

\textbf{Updating Orthogonal Dictionary}: given the sparse coefficients $\textbf{A}^{(0)}$, the sparce coefficients 
$\textbf{A}^{(t)}$ are obtained via solving
\begin{equation}
\begin{split}
\textbf{D}_{i}^{(t+1)}:=
&
\argmin_{\textbf{A}\in\mathbb{R}^{3p^2\times MN}}
\|\mathbf{Y}-[\mathbf{D}_{e};\mathbf{D}_{i}]\mathbf{A}^{(t)}\|_{F}^{2}
\\
&
\text{s.t.}
\quad
\mathbf{D}_{i}^{T}\mathbf{D}_{i} = \mathbf{I}_{r},\ \mathbf{D}_{e}^{T}\mathbf{D}_{i} = \mathbf{0},
\end{split}
\end{equation}

\subsection{Discussion on External Prior and Internal Orthogonal Dictionary Learning}
Until now, we have divided the noisy PGs into multiple internal subspaces. Here we take a deep analysis on how the external NSS prior guide the subspace learning of internal PGs. The help are at least threefold. Firstly, through MAP in (\ref{equ3}), the external prior guides the noisy PGs to be clustered into the correct subspaces. If we cluster the noisy PGs in an automatical way, the subspaces we learned will be highly degraded by the signal dependent noise. Secondly, the guidance of external prior for internal clustering is more efficient than directly clustering the internal noisy PGs. It only needs to calculate the MAP probability via the equation (\ref{equ3}) while the internal clustering via GMM is time-consuming on EM algorithm \cite{em}. Thirdly, due to the correct guidance of external prior, the structual decomposition via SVD of each subspace is more adaptive. This will bring better denoising performance than the methods only using the external information.

@@

\section{The Denoising Algorithm}
\subsection{Adaptive Sparse Coding}
To remove the real noise, we employ the sparse coding framework. And in order to be adaptive to the input image, we employ the internal learned $\mathbf{U}$ of each cluster as an adaptive dictioanry to represent the structural variations of the PGs in that cluster. The \emph{mutual incoherence} $\mu(\mathbf{U})$ \cite{donoho2003optimally}, which is difined as
\begin{equation}
\mu(\mathbf{U})=\max_{i~=j}\frac{|\mathbf{d}_{i}^{T}\mathbf{d}_{j}|}{\|\mathbf{d}_{i}\|_{2}\|\mathbf{d}_{j}\|_{2}}
\end{equation},
is a measure of quality of dictionary. Since the $\mathbf{U}$ is orthonormal, its \emph{mutual incoherence} is naturally $0$ and therefore better than other redundant dictionaries.

\begin{equation}\label{equ5}
\min\nolimits_{\boldsymbol{\upalpha}}\|\mathbf{\overline{y}}_{m}-\mathbf{U}\boldsymbol{\upalpha}\|_{2}^{2}+\sum_{i=1}^{3p^{2}}\lambda_{i}|\boldsymbol{\upalpha}_{i}|.
\end{equation}
The $i$th entry of the regularization parameter $\lambda_{i}$ 
\begin{equation}\label{equ6}
\lambda_{i} = \lambda/(\mathbf{S}_{i}+\varepsilon),
\end{equation}
where $\varepsilon$ is a small positive number to avoid dividing by zero. Since the dictionary $\mathbf{U}$ is orthonormal, it is not difficult to find out that (\ref{equ4}) has a closed-form solution (detailed derivation can be found in the supplementary material):
\begin{equation}\label{equ7}
\setlength{\abovedisplayskip}{3pt}
\setlength{\belowdisplayskip}{4pt}
\hat{\boldsymbol{\upalpha}}= \text{sgn}(\mathbf{U^{T}\overline{y}}_{m})\odot \text{max}(|\mathbf{U^{T}\overline{y}}_{m}|-\mathbf{\Lambda},\mathbf{0}),
\end{equation}
where $\mathbf{\Lambda} = [\lambda_{1},\lambda_{2},...,\lambda_{3p^2}]$ is the vector of regularization parameter and $\text{sgn}(\bullet)$ is the sign function, $\odot$ means element-wise multiplication, and $|\mathbf{U^{T}\overline{y}}_{m}|$ is the absolute value of each entry of vector $|\mathbf{U^{T}\overline{y}}_{m}|$. The closed-form solution makes our weighted sparse coding process very efficient.

\subsection{Fast Patch Group Searching by Integral Image}
The searching of patch groups in images is inefficient if we search non-local similar patches to each local patch. To speed up the searching process and make our proposed method faster, we employ the technique of 'Summed Area Table' \cite{sat} for efficient PG searching. The SAT permits to evaluate the sum of pixel values in rectangular regions of the image with four operations, regardless of the region size. That is to say, we do not need do distance measure for each patch. It was first proposed under the name of summed area table\cite{viola2004robust}

\subsection{Effectively dealing with different noisy images}
For real image denoising, we can perform well on images which have similar noise levels with the training dataset. How can we deal with the real noisy images whose noise levels are higher than the training dataset? The answer is to remove the noise by more iterations. The input image of each iteration is the recovered image of previous iteration. This makes sense since we can still view the recovered image as a real noisy image.

This will also bring a second problem, that how we could automatically terminate the iteration. This can be solved by two methods. One way is to compare the images between two iterations and calculate their difference, the iteration can be terminated if the difference is smaller than a threshold. The other way is to estimate the noise level of the current image and terminate the iterations when the noise level is lower than a preset threshold. We employ the second way and set the threshold as 0.002 in our experiments. In fact, most of our testing images will be denoised well in one iteration.

@@

\subsection{The Overall Algorithm}
With the solution $\hat{\boldsymbol{\upalpha}}$ in (\ref{equ7}), the clean patch in a PG can be estimated as $\hat{\mathbf{x}}_{m}=\mathbf{D}\hat{\boldsymbol{\upalpha}}+\boldsymbol{\upmu}_{y}$. Then the clean image $\hat{\mathbf{x}}$ can be reconstructed by aggregating all the estimated PGs. In practice, we could perform the above denoising procedures for several iterations for better denoising outputs. In iteration $t$, we use the iterative regularization strategy \cite{osher2005iterative} to add back to the recovered image $\hat{\mathbf{x}}^{(t-1)}$ some estimation residual in iteration $t-1$. The proposed denoising algorithm is summarized in Algorithm 1 (Alg. 1).

\begin{table}
\label{alg1}
\begin{tabular}{l}
\hline
\textbf{Alg. 1}: External Guided Internal Dictionary Learning
\\
\hline
\textbf{Input:} Noisy image $\mathbf{y}$, PG-GMM model
\\
1. Initialization: $\hat{\mathbf{x}}^{(0)}=\mathbf{y},\mathbf{y}^{(0)}=\mathbf{y}$;
\\
\textbf{for} $t = 1:IteNum$ \textbf{do}
\\
\quad\textbf{for} each PG $\mathbf{Y}$ \textbf{do}
\\
2.\quad Calculate group mean $\boldsymbol{\upmu}_{y}$ and form PG $\mathbf{\overline{Y}}$;
\\
3.\quad Gaussian component selection via (\ref{equ3});
\\
\quad\textbf{end for}
\\
\quad\textbf{for} each Internal Subspace \textbf{do}
\\
4.\quad Internal Subspace Learning by (\ref{equ4});
\\
5.\quad Recover each patch in all PGs via $\hat{\mathbf{x}}_{m}=\mathbf{D}\hat{\boldsymbol{\upalpha}}+\boldsymbol{\upmu}_{y}$;
\\
\quad\textbf{end for}
\\
6. Aggregate the recovered PGs of all subspaces to form
\\
\quad the recovered image $\hat{\mathbf{x}}^{(t)}$;
\\
\textbf{end for}
\\
\textbf{Output:} The recovered image $\hat{\mathbf{x}}^{(IteNum)}$.\\
\hline
\end{tabular}
\end{table}

>>>>

Though each PG has been projected into its most suitable subspace, the pre-learned subspace is still too general to represent the noisy PG extracted from the real noisy image. That is, the noisy PGs projected into one cluster can still constisted a subspace which is of lower dimensions than the subspace pre-learned from the external PGs. This can be demonstrated by compare the distribution of external PGs and internal PGs in the same clusters. We randomly select one cluster, and collect the celan PGs extracted from external dataset (Kodak 24 images) and the niosy PGs from the testing image. Since the original PGs are of $3p^{2}$ dimensions, we apply PCA to project the PGs into 2 dimensions for better visualization. The results is shown in Figure \ref{fig}, from which we can see clearly that the projected PGs are mainly in a smaller region of the external PGs, which proves that the internal PGs are only consisted a subspace in a lower dimension than the PGs collected from external subspace. To better and adaptively charactering the internal PGs from the testing image, we need learn a more specific dictionary for noisy PGs assigned into each cluster. For notation simplicity, we ignore the index of subspace $k$. The internal PGs $\mathbf{Y}$ form a subspace which can be obtained by singular value decomposition (SVD),
\begin{equation}\label{equ4}
\begin{split}
&\min_{\mathbf{D}_{i}\in \mathbb{R}^{3p^2\times r},\textbf{A}\in\mathbb{R}^{3p^2\times MN}}
\|\mathbf{Y}-[\mathbf{D}_{e}\ \mathbf{D}_{i}]\mathbf{A}\|_{F}^{2}+\lambda\|\mathbf{A}\|_{1}
\\
&
\text{s.t.}
\quad
\mathbf{D}_{i}^{T}\mathbf{D}_{i} = \mathbf{I}_{r},\ \mathbf{D}_{e}^{T}\mathbf{D}_{i} = \mathbf{0},
\end{split}
\end{equation}
The singular vectors capture the statistical structures of NSS variations in natural images, while the singular values in $\mathbf{S}$ represent the significance of these singular vectors. Fig.\ 4 shows the singular vectors for one Gaussian component.

@@

\textbf{Updating Sparse Coefficients}: given the initialization orthogonal dictioanry $\textbf{D}_{i}^{(t)}$, the sparce coefficients 
$\textbf{A}^{(t)}$ are obtained via solving
\begin{equation}\label{equ5}
\textbf{A}^{(t)}:=\argmin_{\textbf{A}\in\mathbb{R}^{3p^2\times MN}}
\|\mathbf{Y}-[\mathbf{D}_{e}\ \mathbf{D}_{i}^{(t)}]\mathbf{A}\|_{F}^{2}+\lambda\|\mathbf{A}\|_{1}.
\end{equation}
This problem has closed-form solution by $\mathbf{A}^{*}=T_{\lambda}(\mathbf{\hat{D}}^{T}\mathbf{Y})$, 
where $T_{\lambda}(\mathbf{A}) = \text{sgn}(\mathbf{A})\odot\max(\mathbf{A},\lambda)$ is a soft-thresholding function.

\textbf{Updating Orthogonal Dictionary}: given the sparse coefficients $\textbf{A}^{(0)}$, the sparce coefficients 
$\textbf{A}^{(t)}$ are obtained via solving
\begin{equation}\label{equ6} 
\begin{split}
\textbf{D}_{i}^{(t+1)}:=
&
\argmin_{\textbf{D}_{i}\in\mathbb{R}^{3p^2\times r}}
\|\mathbf{Y}-[\mathbf{D}_{e}\ \mathbf{D}_{i}]\mathbf{A}^{(t)}\|_{F}^{2}
\\
&
\text{s.t.}
\quad
\mathbf{D}_{i}^{T}\mathbf{D}_{i} = \mathbf{I}_{r},\ \mathbf{D}_{e}^{T}\mathbf{D}_{i} = \mathbf{0},
\end{split}
\end{equation}
Dividing the sparse coefficients $\mathbf{A}=[\mathbf{A}_{e}^{T}\ \mathbf{A}_{i}^{T}]^{T}$, where $\mathbf{A}_{e}$ and $\mathbf{A}_{i}$ denote the coefficients over external and internal dictionary $\mathbf{D}_{e}$ and $\mathbf{D}_{i}$. According to the Proposition 2.2 in \cite{bao2013fast}, the problem (\ref{equ6}) 
%\begin{equation}
%\begin{split}
%\min_{\textbf{D}_{i}\in\mathbb{R}^{3p^2\times r}}
%&\|\mathbf{Y}-\mathbf{D}_{e}\mathbf{A}_{e}+\mathbf{D}_{i}\mathbf{A}_{i}\|_{F}^{2}
%\\
%&
%\text{s.t.}
%\quad
%\mathbf{D}_{i}^{T}\mathbf{D}_{i} = \mathbf{I}_{r},\ \mathbf{D}_{e}^{T}\mathbf{D}_{i} = \mathbf{0},
%\end{split}
%\end{equation}
has a closed-form solution $\mathbf{D}_{i}^{*}=\mathbf{U}\mathbf{V}^{T}$, where $\mathbf{U}$ and $\mathbf{V}$ are
the orthogonal matrices obtained by the following SVD
\begin{equation}\label{equ7}
(\mathbf{I}-\mathbf{D}_{e}\mathbf{D}_{e}^{T})\mathbf{Y}\mathbf{A}_{i}^{T}
=
\mathbf{U}\Sigma\mathbf{V}^{T}
\end{equation}
With these solutions, the final obtained dictionary $\mathbf{D} = [\mathbf{D}_{e}\ \mathbf{D}_{i}]$ are orthogonal ictionary. This can be proved by the following equation
\begin{equation}\label{equ8}
\mathbf{D}^{T}\mathbf{D} = 
\left(\begin{array}{c}
\mathbf{D}_{e}^{T}
\\
\mathbf{D}_{i}^{T}
\end{array}\right)
(\mathbf{D}_{e}\ \mathbf{D}_{i})
=
\left(\begin{array}{cc}
\mathbf{D}_{e}^{T}\mathbf{D}_{e}\ \mathbf{D}_{e}^{T}\mathbf{D}_{i}
\\
\mathbf{D}_{i}^{T}\mathbf{D}_{e}\ \mathbf{D}_{i}^{T}\mathbf{D}_{i}
\end{array}\right)
=
\mathbf{I}
\end{equation}

\subsection{Discussion on External Prior and Internal Orthogonal Dictionary Learning}
Until now, we have divided the noisy PGs into multiple internal subspaces. Here we take a deep analysis on how the external NSS prior guide the subspace learning of internal PGs. The help are at least threefold. Firstly, through MAP in (\ref{equ3}), the external prior guides the noisy PGs to be clustered into the correct subspaces. If we cluster the noisy PGs in an automatical way, the subspaces we learned will be highly degraded by the signal dependent noise. Secondly, the guidance of external prior for internal clustering is more efficient than directly clustering the internal noisy PGs. It only needs to calculate the MAP probability via the equation (\ref{equ3}) while the internal clustering via GMM is time-consuming on EM algorithm \cite{em}. Thirdly, due to the correct guidance of external prior, the structual decomposition via SVD of each subspace is more adaptive. This will bring better denoising performance than the methods only using the external information. The \emph{mutual incoherence} $\mu(\mathbf{U})$ \cite{donoho2003optimally}, which is difined as
\begin{equation}\label{equ9}
\mu(\mathbf{U})=\max_{i~=j}\frac{|\mathbf{d}_{i}^{T}\mathbf{d}_{j}|}{\|\mathbf{d}_{i}\|_{2}\|\mathbf{d}_{j}\|_{2}}
\end{equation},
is a measure of quality of dictionary.

@@

\section{The Denoising Algorithm}
\subsection{Fast Patch Group Searching by Integral Image}
The searching of patch groups in images is inefficient if we search non-local similar patches to each local patch. To speed up the searching process and make our proposed method faster, we employ the technique of 'Summed Area Table' \cite{sat} for efficient PG searching. The SAT permits to evaluate the sum of pixel values in rectangular regions of the image with four operations, regardless of the region size. That is to say, we do not need do distance measure for each patch. It was first proposed under the name of summed area table\cite{viola2004robust}

\subsection{Prior Weights for Sparse Coding}
To remove the real noise, we employ the sparse coding framework. And in order to be adaptive to the input image, we employ the internal learned $\mathbf{U}$ of each cluster as an adaptive dictioanry to represent the structural variations of the PGs in that cluster. Since the $\mathbf{U}$ is orthonormal, its \emph{mutual incoherence} is naturally $0$ and therefore better than other redundant dictionaries.

\begin{equation}\label{equ10}
\min\nolimits_{\boldsymbol{\upalpha}}\|\mathbf{\overline{y}}_{m}-\mathbf{U}\boldsymbol{\upalpha}\|_{2}^{2}+\sum_{i=1}^{3p^{2}}\lambda_{i}|\boldsymbol{\upalpha}_{i}|.
\end{equation}
The $i$th entry of the regularization parameter $\lambda_{i}$ 
\begin{equation}\label{equ11}
\lambda_{i} = \lambda/(\mathbf{S}_{i}+\varepsilon),
\end{equation}
where $\varepsilon$ is a small positive number to avoid dividing by zero. Since the dictionary $\mathbf{U}$ is orthonormal, it is not difficult to find out that (\ref{equ4}) has a closed-form solution (detailed derivation can be found in the supplementary material):
\begin{equation}\label{equ12}
\setlength{\abovedisplayskip}{3pt}
\setlength{\belowdisplayskip}{4pt}
\hat{\boldsymbol{\upalpha}}= \text{sgn}(\mathbf{U^{T}\overline{y}}_{m})\odot \text{max}(|\mathbf{U^{T}\overline{y}}_{m}|-\mathbf{\Lambda},\mathbf{0}),
\end{equation}
where $\mathbf{\Lambda} = [\lambda_{1},\lambda_{2},...,\lambda_{3p^2}]$ is the vector of regularization parameter and $\text{sgn}(\bullet)$ is the sign function, $\odot$ means element-wise multiplication, and $|\mathbf{U^{T}\overline{y}}_{m}|$ is the absolute value of each entry of vector $|\mathbf{U^{T}\overline{y}}_{m}|$. The closed-form solution makes our weighted sparse coding process very efficient.

%\subsection{Effectively dealing with different noisy images}
%For real image denoising, we can perform well on images which have similar noise levels with the training dataset. How can we deal with the real noisy images whose noise levels are higher than the training dataset? The answer is to remove the noise by more iterations. The input image of each iteration is the recovered image of previous iteration. This makes sense since we can still view the recovered image as a real noisy image.

%This will also bring a second problem, that how we could automatically terminate the iteration. This can be solved by two methods. One way is to compare the images between two iterations and calculate their difference, the iteration can be terminated if the difference is smaller than a threshold. The other way is to estimate the noise level of the current image and terminate the iterations when the noise level is lower than a preset threshold. We employ the second way and set the threshold as 0.002 in our experiments. In fact, most of our testing images will be denoised well in one iteration.

@@

\subsection{The Overall Algorithm}
With the solution $\hat{\boldsymbol{\upalpha}}$ in (\ref{equ7}), the clean patch in a PG can be estimated as $\hat{\mathbf{x}}_{m}=\mathbf{D}\hat{\boldsymbol{\upalpha}}+\boldsymbol{\upmu}_{y}$. Then the clean image $\hat{\mathbf{x}}$ can be reconstructed by aggregating all the estimated PGs. In practice, we could perform the above denoising procedures for several iterations for better denoising outputs. In iteration $t$, we use the iterative regularization strategy \cite{osher2005iterative} to add back to the recovered image $\hat{\mathbf{x}}^{(t-1)}$ some estimation residual in iteration $t-1$. The proposed denoising algorithm is summarized in Algorithm 1 (Alg. 1).
\begin{table}
\label{alg1}
\begin{tabular}{l}
\hline
\textbf{Alg. 1}: External Prior Guided Internal Orthogonal 
\\
\quad \quad \quad Dictionary Learning for Denoising
\\
\hline
\textbf{Input:} Noisy image $\mathbf{y}$, PG-GMM model
\\
1. Initialization: $\hat{\mathbf{x}}^{(0)}=\mathbf{y},\mathbf{y}^{(0)}=\mathbf{y}$;
\\
\textbf{for} $t = 1:IteNum$ \textbf{do}
\\
\quad\textbf{for} each PG $\mathbf{Y}$ \textbf{do}
\\
2.\quad Calculate group mean $\boldsymbol{\upmu}_{y}$ and form PG $\mathbf{\overline{Y}}$;
\\
3.\quad Gaussian component selection via (\ref{equ3});
\\
\quad\textbf{end for}
\\
\quad\textbf{for} each Internal Subspace \textbf{do}
\\
4.\quad Internal Subspace Learning by (\ref{equ4});
\\
5.\quad Recover each patch in all PGs via $\hat{\mathbf{x}}_{m}=\mathbf{D}\hat{\boldsymbol{\upalpha}}+\boldsymbol{\upmu}_{y}$;
\\
\quad\textbf{end for}
\\
6. Aggregate the recovered PGs of all subspaces to form
\\
\quad the recovered image $\hat{\mathbf{x}}^{(t)}$;
\\
\textbf{end for}
\\
\textbf{Output:} The recovered image $\hat{\mathbf{x}}^{(IteNum)}$.\\
\hline
\end{tabular}
\end{table}
